Are you normal? A new projection pursuit index to assess a sample against a multivariate null distribution

### 2024-07-01 paper.qmd
example titles
add theorem/proof structure

---
title: "Are you normal? A new projection pursuit index to assess a sample against a multivariate null distribution"
author: "Annalisa Calvi, Ursula Laa, Dianne Cook"
format:
  jasa-pdf:
    keep-tex: true    
bibliography: bibliography.bib  
header-includes: | 
  \usepackage{amsmath}
  \usepackage{float}
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \usepackage{bm}
  \def\tightlist{}
  \usepackage{setspace}
  \newcommand\pD{$p\text{-}D$}
  \newcommand\kD{$k\text{-}D$}
  \newcommand\dD{$d\text{-}D$}
  \newcommand\gD{$2\text{-}D$}
---

@@

The projection of an ellipsoid onto \gD{} is an ellipse, where the curve of the ellipse is defined through the set of points $\mathbfit{x}$ for which the gradient is parallel to the projection plane. That is, the curve 
consists of $\mathbfit{x}$ satisfying

@@

We call points in the projection that are on the curve $\mathbfit{y}$, so that $\mathbfit{y} = \mathbfit{x}P$ for $\mathbfit{x}$ on the \pD{} curve. Then $\mathbfit{y} - \mathbfit{\mu}_p = (\mathbfit{x} - \mathbfit{\mu})P = \mathbfit{s} P^T \Sigma P$, where $\mathbfit{\mu}_p = \mathbfit{\mu} P$ is the projected mean. We then substitute $(\mathbfit{y} - \mathbfit{\mu}_p) (P^T \Sigma P)^{-1}$ for $\mathbfit{s}$ in the previous equation. From this we can compute the analog equation for the projection as

$$(\mathbfit{y} - \mathbfit{\mu}_p)(P^T \Sigma P)^{-1}(\mathbfit{y} - \mathbfit{\mu}_p)^T = c^2.$$
This means the matrix $(P^T \Sigma P)^{-1}$ is defining the ellipse in the \gD{} projection. In general $c$ could be any constant, but typically we would select it as a quantile of the $\chi^2$ distribution, so that the size of the ellipse corresponds to a selected probability.

@@

### Robust statistics

### Physics?

>>>>

---
title: "Are you normal? A new projection pursuit index to assess a sample against a multivariate null distribution"
author: "Annalisa Calvi, Ursula Laa, Dianne Cook"
format:
  jasa-pdf:
    keep-tex: true    
bibliography: bibliography.bib  
header-includes: | 
  \usepackage{amsmath}
  \usepackage{amsthm}
  \usepackage{float}
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \usepackage{bm}
  \def\tightlist{}
  \usepackage{setspace}
  \newcommand\pD{$p\text{-}D$}
  \newcommand\kD{$k\text{-}D$}
  \newcommand\dD{$d\text{-}D$}
  \newcommand\gD{$2\text{-}D$}
  \newtheorem*{theorem}{Theorem}
---

@@

\begin{theorem}
  The projection of this \pD{} ellipsoid in \gD{} has the equation
$$(\mathbfit{y} - \mathbfit{\mu}_p)(P^T \Sigma P)^{-1}(\mathbfit{y} - \mathbfit{\mu}_p)^T = c^2.$$
\end{theorem}
\begin{proof}
The projection of an ellipsoid onto \gD{} is an ellipse, where the curve of the ellipse is defined through the set of points $\mathbfit{x}$ for which the gradient is parallel to the projection plane. That is, the curve 
consists of $\mathbfit{x}$ satisfying

@@

We call points in the projection that are on the curve $\mathbfit{y}$, so that $\mathbfit{y} = \mathbfit{x}P$ for $\mathbfit{x}$ on the \pD{} curve. Then $\mathbfit{y} - \mathbfit{\mu}_p = (\mathbfit{x} - \mathbfit{\mu})P = \mathbfit{s} P^T \Sigma P$, where $\mathbfit{\mu}_p = \mathbfit{\mu} P$ is the projected mean. We then substitute $(\mathbfit{y} - \mathbfit{\mu}_p) (P^T \Sigma P)^{-1}$ for $\mathbfit{s}$ in the equation \(c^2 = \mathbfit{s} P^T \Sigma P \mathbfit{s}^T\). From this we can compute the analog equation for the projection as

$$(\mathbfit{y} - \mathbfit{\mu}_p)(P^T \Sigma P)^{-1}(\mathbfit{y} - \mathbfit{\mu}_p)^T = c^2$$

as claimed.
\end{proof}

This means the matrix $(P^T \Sigma P)^{-1}$ is defining the ellipse in the \gD{} projection. In general $c$ could be any constant, but typically we would select it as a quantile of the $\chi^2$ distribution, so that the size of the ellipse corresponds to a selected probability.

@@

### Robust statistics: weather extremes
